\LARGE \bf Exploiting Non-Slip Wall Contacts \\ to Position Two Particles Using The Same Control Input

### 2018-05-01 journalWallFriction/Experiment.tex
Added a lot of new pictures, edited a little bit.

 The magnetic manipulation system has two pairs of electromagnetic coils each with iron cores at their centers, and arranged orthogonal to each other. The iron core at the center of each coil concentrates the magnetic field towards the workspace. An Arduino and four SyRen regenerative motor drivers were used for control inputs to the coils. Finally, a FOculus F0134SB 659 x 494 pixel camera was attached to the top of the system focusing on the workspace which was back-lit by a 15w LED light strip.

To obtain experimental data, the test samples which comprised of the phantom intestine model and the bovine cross section, were placed in laser cut acrylic discs and then immersed in corn syrup. Corn syrup was used to have the best viscosity(12000 cP) for the experiments. The velocities of particles immersed in it were dampened enough to control their movements with ease. Spherical $0.5 mm$ magnets(supermagnetman #SP0100-50) were used as our particles.

@@

The intestine phantom model was used as the first test field of the project and was made to mimic the geometry of an intestine and its villi. The model consists of a circular ring with an outer diameter of $\SI{50}{\milli\metre}$, an inner diameter of $\SI{46}{\milli\metre}$, and 60 $2 mm$ long protrusions on its inner surface cut out of $6 mm$ thick acrylic to model the geometry of intestinal villi. Fig.~\ref{fig:3dPrinted} shows an experiment with starting and ending positions drawn with marker on the workspace.

@@

\begin{figure*}\label{fig:3dPrinted}
\centering
\vspace{1.5em}
%\begin{overpic}[width=\columnwidth]{firstImage.jpg}\end{overpic}
\begin{overpic}[width=2\columnwidth]{3dexperiment.pdf}\end{overpic}
\caption{\label{fig:story}
Positioning particles that receive the same control inputs, but cannot move while a control input pushes them into a boundary.
} \vspace{-1em}
\end{figure*}

>>>>

 The magnetic manipulation system has two pairs of electromagnetic coils each with iron cores at their centers, and arranged orthogonal to each other. The iron core at the center of each coil concentrates the magnetic field towards the workspace. An Arduino and four SyRen regenerative motor drivers were used for control inputs to the coils. Finally, a FOculus F0134SB 659 x 494 pixel camera was attached to the top of the system focusing on the workspace which was back-lit by a 15$w$ LED light strip.

To obtain experimental data, the test samples which comprised of the phantom intestine model and the bovine cross section, were placed in laser cut acrylic discs and then immersed in corn syrup. Corn syrup was used to have the best viscosity(12000 cP) for the experiments. The velocities of particles immersed in it were dampened enough to control their movements with ease. Spherical $\SI{0.5}{\milli\metre}$ magnets (supermagnetman \#SP0100-50) were used as our particles.

@@

The intestine phantom model was used as the first test field of the project and was made to mimic the geometry of an intestine and its villi. The model consists of a circular ring with an outer diameter of $\SI{50}{\milli\metre}$, an inner diameter of $\SI{46}{\milli\metre}$, and 60 $\SI{2}{\milli\metre}$ long protrusions on its inner surface cut out of $\SI{6}{\milli\metre}$ thick acrylic to model the geometry of intestinal villi. Fig.~\ref{fig:3dPrinted} shows an experiment with starting and ending positions drawn with marker on the workspace.

@@

\begin{figure*}\label{fig:3dPrinted}
\centering
\vspace{1.5em}
%\begin{overpic}[width=\columnwidth]{firstImage.jpg}\end{overpic}
\begin{overpic}[width=2\columnwidth]{3dexperiment.pdf}\end{overpic}
\\
\vspace{1em}
\begin{overpic}[width=2\columnwidth]{realTripe.pdf}\end{overpic}
\caption{\label{fig:story}
Positioning particles that receive the same control inputs, but cannot move while a control input pushes them into a boundary.
} \vspace{-1em}
\end{figure*}

### 2018-05-01 journalWallFriction/SwarmShapeControl.tex
Edited a lot of parts of the paper
Edited algorithms parts.
Added a lot of new pictures, edited a little bit.

%\usepackage{wrapfig}
\usepackage{overpic}
%\usepackage{subfigure}
%\usepackage{textcomp}
\graphicspath{{./pictures/pdf/},{./pictures/ps/},{./pictures/png/},{./pictures/jpg/}}
\usepackage{breqn} %for breaking equations automatically
\usepackage[ruled]{algorithm}
\usepackage{algpseudocode}
%\usepackage{algorithmic}
\usepackage{multirow}
\usepackage{todonotes}
\usepackage{authblk}

%\newcommand{\todo}[1]{\vspace{5 mm}\par \noindent \framebox{\begin{minipage}[c]{0.98 \columnwidth} \ttfamily\flushleft \textcolor{red}{#1}\end{minipage}}\vspace{5 mm}\par}
% uncomment this to hide all red todos
%\renewcommand{\todo}{}

@@

Steered particles offer a method for targeted therapy, interventions, and drug delivery in regions inaccessible by large robots.
Magnetic actuation has the benefits of requiring no tethers, being able to operate from a distance, and in some cases allows imaging for feedback (e.g. MRI).
 This paper investigates control with uniform magnetic gradients (the same force is applied everywhere in the workspace).
Given three orthogonal magnetic fields, steering one particle in 3D is trivial. 
Adding additional particles to steer makes the system underactuated because there are more states than control inputs. 
However, the walls of in vivo and artificial environments often have surface roughness such that the particles do not move unless actuation pulls them away from the wall.
In previous works, we showed that the individual 2D position of two particles is controllable in a square workspace with non-slip wall contact. 
Because in vivo environments are usually not square, this work extends the previous work to convex workspaces including circles and 3D positioning. 
This paper also implements the algorithms using a hardware setup inspired by gastrointestinal tract.

>>>>

%\usepackage{wrapfig}
\usepackage{overpic}
%\usepackage{subfigure}
%\usepackage{textcomp}
\graphicspath{{./pictures/pdf/},{./pictures/ps/},{./pictures/png/},{./pictures/jpg/}}
\usepackage{breqn} %for breaking equations automatically
\usepackage[ruled]{algorithm}
\usepackage{algpseudocode}
%\usepackage{algorithmic}
\usepackage{multirow}
\usepackage{todonotes}
\usepackage{authblk}
\usepackage[per-mode=symbol, detect-weight=true, binary-units=true]{siunitx}
\usepackage{rotating}
%\newcommand{\todo}[1]{\vspace{5 mm}\par \noindent \framebox{\begin{minipage}[c]{0.98 \columnwidth} \ttfamily\flushleft \textcolor{red}{#1}\end{minipage}}\vspace{5 mm}\par}
% uncomment this to hide all red todos
%\renewcommand{\todo}{}

@@

Steered particles offer a method for targeted therapy, interventions, and drug delivery in regions inaccessible by large robots.
Magnetic actuation has the benefits of requiring no tethers, being able to operate from a distance, and in some cases allows imaging for feedback (e.g. MRI).
 This paper investigates particle control with uniform magnetic gradients (the same force is applied everywhere in the workspace).
Given three orthogonal magnetic fields, steering one particle in 3D is trivial. 
Adding additional particles to steer makes the system underactuated because there are more states than control inputs. 
However, the walls of in vivo and artificial environments often have surface roughness such that the particles do not move unless actuation pulls them away from the wall.
In previous works, we showed that the individual 2D position of two particles is controllable in a square workspace with non-slip wall contact. 
Because in vivo environments are usually not square, this paper extends the previous work to convex workspaces including circles and 3D positioning. 
This paper also implements the algorithms using a hardware setup inspired by the gastrointestinal tract.

### 2018-05-01 journalWallFriction/intro.tex
Edited a lot of parts of the paper
Added a lot of new pictures, edited a little bit.

The small size of these robots makes it difficult to perform onboard computation.  Instead, these robots are often controlled by a broadcast signal. 
 The tiny robots themselves are often just rigid bodies, and it may be more accurate to define the \emph{system}, consisting of particles, a uniform control field, and sensing, as the robot.
Such systems are severely underactuated, having 2 degrees of freedom in the shared control input, but $2n$ degrees of freedom for the particle swarm.
 Techniques are needed that can handle this underactuation. 
 In previous work, we showed that the 2D position of each particle in such a swarm is controllable if the workspace contains a single obstacle the size of one particle.

@@

The paper is arranged as follows. 
After a review of recent related work in Sec.  \ref{sec:RelatedWork},
 %Sec. \ref{subsec:FluidInTank} provides analytical position control results of stable configurations in two canonical workspaces with frictionless walls.  These results are limited in the set of shapes that can be generated.  To extend the range of possible shapes,
  Sec.  \ref{subsec:WallFriction} introduces a  model for boundary interaction.  
We provide a shortest-path algorithm to arbitrarily position two robots in Sec.  \ref{sec:PostionControl2Robots}.
Sec.  \ref{sec:simulation} describes implementations of the algorithms in simulation and  Sec.  \ref{sec:expResults} describes hardware experiments, as shown in Fig.~\ref{fig:IntroPic}. 
 We end with directions for future research in Sec.  \ref{sec:conclusion}.

>>>>

The small size of these robots makes it difficult to perform onboard computation.  Instead, these robots are often controlled by a broadcast signal. 
 The tiny robots themselves are often just rigid bodies, and it may be more accurate to define the \emph{system}, consisting of particles, a uniform control field, and sensing, as the robot.
Such systems are severely underactuated, having 2 degrees of freedom in the shared control input, but $2n$ degrees of freedom for the $n$-particle swarm.
 Techniques are needed that can handle this underactuation. 
 In previous work, we showed that the 2D position of each particle in such a swarm is controllable if the workspace contains a single obstacle the size of one particle \cite{AaronManipulation2013}.

@@

The paper is arranged as follows. 
After a review of recent related work in Sec.  \ref{sec:RelatedWork},
 %Sec. \ref{subsec:FluidInTank} provides analytical position control results of stable configurations in two canonical workspaces with frictionless walls.  These results are limited in the set of shapes that can be generated.  To extend the range of possible shapes,
  Sec.  \ref{subsec:WallFriction} introduces a  model for boundary interaction.  
We provide a shortest-path algorithm to arbitrarily position two particles in Sec.  \ref{sec:PostionControl2Robots}.
Sec.  \ref{sec:simulation} describes implementations of the algorithms in simulation and  Sec.  \ref{sec:expResults} describes hardware experiments, as shown in Fig.~\ref{fig:IntroPic}. 
 We end with directions for future research in Sec.  \ref{sec:conclusion}.

### 2018-05-01 journalWallFriction/theory.tex
Edited algorithms parts.
Added a lot of new pictures, edited a little bit.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Theory}
\label{sec:theory}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
If a control input causes the particles to collide with obstacles at different times, inverting the control input does not undo the action. Due to this lack of time-reversibility, techniques that require a bidirectional graph, e.g. PRM and RRT* are not suited \cite{lavalle2006planning,kavraki1996probabilistic}. Instead, this paper employs a graph search. This section starts with a boundary interaction model in subsection \ref{subsec:WallFriction}.

@@

In the absence of obstacles uniform inputs move a swarm identically.  
Independent control requires breaking this symmetry. 
The following sections examine using non-slip boundary contacts to break the symmetry caused by uniform inputs.

@@

These system dynamics represent particle swarms in low-Reynolds number environments, where viscosity dominates inertial forces and so velocity is proportional to input force~\cite{Purcell1977}. 
 In this regime, the input force command $\mathbf{u}(t)$ controls the velocity of the particles.  
  The same model can be generalized to particles moved by fluid flow where the vector direction of fluid flow $\mathbf{u}(t)$ controls the velocity of particles, or for a swarm of particles that move at a constant speed in a direction specified by a uniform input $\mathbf{u}(t)$~\cite{Rubenstein2012}.
  As in our model, fluid flowing in a pipe has zero velocity along the boundary. Similar mechanical systems exist at larger scales, e.g. all tumblers of a combination look move uniformly unless obstructed by an obstacle.
 Our control problem is to design the control inputs $\mathbf{u}(t)$ to make all $n$ particles achieve a task.

@@

 \subsection{Shortest Path in Unit Disk that Intersects Circumference}\label{subsec:circular}
 The shortest path between two points in the unit disk that reflects off the circumference is composed of two straight line segments, as shown in Fig.~\ref{fig:shortestpath}. The problem can be simplified by choosing the coordinate system carefully. We define the $x$ axis along the position of the starting point: $S=(s,0)$, and define the point of intersection by the angle $\theta$ from the $x$ axis: $P=(\cos \theta,\sin \theta)$. Define the final point $E$ by a radius $e$ and angle $\beta$: $E=e(\cos \beta,\sin \beta)$. Then the length of the two line segments is 
 \begin{equation}
\scriptsize{ \sqrt{(s-\cos \theta)^2+\sin^2 \theta}+\sqrt{(e \sin \beta-\cos \theta)^2+(e \sin \beta-\sin \theta)^2}},
 \end{equation}
 which is minimized by choosing an appropriate $\theta$ value. This equation can be simplified to

@@

\begin{figure}
\centering
\renewcommand{\figwid}{\columnwidth}
{\begin{overpic}[width =\figwid]{shortestpath.pdf}
\end{overpic}
}
\caption{\label{fig:shortestpath}{The shortest path between two points (blue square) to (blue ellipse) in the unit disk that intersects the circumference. The path length as a function of intersection point, $(\cos\theta,\sin\theta)$ is shown at right. See \cite{shortestPathMathematica}.}
%\vspace{-1em}
}
\end{figure}

@@

 Simplifying this system of equations results in $s=e \csc \theta (s \sin(2 \theta-\beta)+\sin(\beta-\theta))$. Solving this last equation results in a quartic solution that has a closed-form solution with four roots, each of which can be either a clockwise or a counterclockwise rotation $\theta$, depending on the sign of $\beta$, with $-\pi\leq\beta\leq\pi$. We evaluate each and select the solution that results in the shortest length path. This optimal path satisfies the law of reflection off the unit circle, with angle of incidence equal to angle of reflection. For an interactive Mathematica demonstration of this shortest path, see \cite{shortestPathMathematica}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Theory}
\label{sec:theory}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
If a control input causes the particles to collide with obstacles at different times, inverting the control input does not undo the action. Due to this lack of time-reversibility, techniques that require a bidirectional graph, e.g. PRM and RRT* are not suitable \cite{lavalle2006planning,kavraki1996probabilistic}. Instead, this paper employs a graph search. This section starts with a boundary interaction model in subsection \ref{subsec:WallFriction}.

@@

In the absence of obstacles, uniform inputs move a swarm identically.  
Independent control requires breaking this symmetry. 
The following sections examine using non-slip boundary contacts to break the symmetry caused by uniform inputs.

@@

These system dynamics represent particle swarms in low-Reynolds number environments, where viscosity dominates inertial forces and so velocity is proportional to input force~\cite{Purcell1977}. 
 In this regime, the input force command $\mathbf{u}(t)$ controls the velocity of the particles.  
  The same model can be generalized to particles moved by fluid flow where the vector direction of fluid flow $\mathbf{u}(t)$ controls the velocity of particles, or for a swarm of particles that move at a constant speed in a direction specified by a uniform input $\mathbf{u}(t)$~\cite{Rubenstein2012}.
  As in our model, fluid flowing in a pipe has zero velocity along the boundary. Similar mechanical systems exist at larger scales, e.g. all tumblers of a combination lock move uniformly unless obstructed by an obstacle.
 Our control problem is to design the control inputs $\mathbf{u}(t)$ to make all $n$ particles achieve a task.

@@

 \subsection{Shortest Path in Unit Disk that Intersects Circumference}\label{subsec:circular}
 The shortest path between two points in the unit disk that intersects the circumference is composed of two straight line segments, as shown in Fig.~\ref{fig:shortestpath}. The problem can be simplified by choosing the coordinate system carefully. We define the $x$ axis along the position of the starting point: $S=(s,0)$, and define the point of intersection by the angle $\theta$ from the $x$ axis: $P=(\cos \theta,\sin \theta)$. Define the final point $E$ by a radius $e$ and angle $\beta$: $E=e(\cos \beta,\sin \beta)$. Then the length of the two line segments is 
 \begin{equation}
\scriptsize{ \sqrt{(s-\cos \theta)^2+\sin^2 \theta}+\sqrt{(e \sin \beta-\cos \theta)^2+(e \sin \beta-\sin \theta)^2}},
 \end{equation}
 which is minimized by choosing an appropriate $\theta$ value. This equation can be simplified to

@@

\begin{figure}
\centering
\renewcommand{\figwid}{\columnwidth}
{\begin{overpic}[width =\figwid]{shortestpath.pdf}
\end{overpic}
}
\caption{\label{fig:shortestpath}{The shortest path between two points (blue square) to (blue ellipse) in the unit disk that intersects the circumference. The path length as a function of intersection point, $(\cos\theta,\sin\theta)$ is shown at right. See \cite{BeckerShortestPath}.}
%\vspace{-1em}
}
\end{figure}

@@

 Simplifying this system of equations results in $s=e \csc \theta (s \sin(2 \theta-\beta)+\sin(\beta-\theta))$. Solving this last equation results in a quartic solution that has a closed-form solution with four roots, each of which can be either a clockwise or a counterclockwise rotation $\theta$, depending on the sign of $\beta$, with $-\pi\leq\beta\leq\pi$. We evaluate each and select the solution that results in the shortest length path. This optimal path satisfies the law of reflection off the unit circle, with angle of incidence equal to angle of reflection. For an interactive Mathematica demonstration of this shortest path, see \cite{BeckerShortestPath}.
